---
title: "Assignment_1"
author: "Logan Luchs"
date: "2025-02-11"
output: html_document
---


1. You type the R Code in the upper left box of R Studio, 
and the box below shows the output of the code that is ran. 
The plots show up in the plots tab of the bottom right hand box.  T
he global environment is all of the objects or data that is stored in your session.
If you store a plot in X it will show up as X in the global environment window. 
You can get help with a function by doing the ?func comamand or by looking in the help tab of the bottom right hand box
An R Package is a collection of functions or commands that can be ran in R.
You can find installed packages by looking in the packages tab (if they are loaded they will have a check mark next to them).
The working directory is the same place where your project has been created. If you go to the files tab you can find where the working directory is by default.
You can set your working directory differently and the project file should move with iterators
A relative file path is in realation to your project file, but absolute file path is from your laptop. Eg; //users/logan/...

2. First create a github account. Create a new repository with a readme file and a liscence of your choosing.
Next download git for windows so that it can be compatible for R.
Install Git and then open RStudio.
Click Open new project.
Click Version Control.
Insert the URL of your newly created repository on github
The rest of that should autopopulate with what you created on github.
Hit create Project.
https://github.com/Logz1n/Reproducibility_Class
is the one I made for the class assignments

3. A vector is a variable with multiple values stored into it eg; x <- 1,2,3,4,5
A dataframe is a table of data with the rows having observations and columns are variables that you are interested in.
Matrices are 2 dimensional data frames that can be made with similar data types. Math is typically done for this.
Statistics are also reported in this format.

```{r}
# Create a vector z with values from 1 to 200
z <- 1:200

# Calculate the mean of vector z
mean(z)

# Calculate the standard deviation of vector z
sd(z)

# Create a logical vector zlog where values in z greater than 30 are labeled "TRUE", otherwise "FALSE"
zlog <- ifelse(z > 30, "TRUE", "FALSE")

# Create a data frame zdf with columns z and zlog
zdf <- data.frame(z, zlog)

# Rename the columns of zdf to "zvec" and "zlogic"
colnames(zdf) <- c("zvec", "zlogic")

# Add a new column zsquared to zdf, which is the square of the values in zvec
zdf$zsquared <- zdf$zvec^2

# Create a subset zdf1 of zdf where zsquared is greater than 10
zdf1 <- zdf[zdf$zsquared > 10,]

# Further subset zdf1 where zsquared is less than 100
zdf1[zdf1$zsquared < 100,]  

# Create another subset zdf2 of zdf where zsquared is greater than 10
zdf2  <- subset(zdf, zsquared > 10)

# Further subset zdf2 where zsquared is less than 100
subset(zdf2, zsquared < 100)

# Access the 26th row of zdf
zdf[26,]

# Access the value in the 180th row and zsquared of zdf
zdf[180,3]
```

```{r}
Tips <- read.csv("Assignment 1/TipsR.csv", sep = ",")
#The file was read correctly because when I looked at the file there was a delimiter seperating the . from the rest of the columns. When looked at in R, the it read the same exact way. I also double made sure by specifying what the delimiter was instead of using the default code.












```

