---
title: "Assignment_6"
author: "Logan Luchs"
date: "2025-03-27"
output:
  md_document:
    variant: gfm
  word_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
  pdf_document:
---
[Link to my Github](https://github.com/Logz1n/Reproducibility_Class)

#1 Writing your own functions and iterations allows other people to use your code for other things. Therefore aiding reproducibility of your code.

#2 for loops start with an initialization such as for (i in X){, then proceed with the function until everything in x has been looped over. A function is placed after the open bracket. For example for(i in x) {logx <- log(i). To return the results print() can be used. and example would be print(logx). After that the function is closed with the closed bracket }. The total code for the example would look like this:
##for(i in x){
##logx <- log(i)
##print(logx)
##}

# Loading libraries and color pallettes
```{r Loading Libraries and Pallettes, message=FALSE, warning=FALSE} 
#Loading Libraries
library(ggplot2) 
library(knitr)
library(readr)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(markdown)
library(drc)

#Colorblind pallette
cbbPalette <- c("#56B4E9", "#009E73", "#F0E442",
"#000000", "#D55E00", "#CC79A7", "#E69F00","#0072B2" ) #loading a color pallette

```

#3. Reading in Data
```{r Reading in Dataset}
Cities <- read.csv("Cities.csv", na = "na" ) #loading in the data so that R understands na is na so the column is numeric

AuburnLatLong <- Cities %>%
  subset(city == "Auburn")
#subsetting to get data for the second city
Cities2 <- Cities %>%
  mutate(city1 = city, city2 = "Auburn", lat2 = AuburnLatLong$lat, lon2 = AuburnLatLong$long)
#Create a column with the second city metrics, since only auburn is wanted
```

#4.The Haversine function
```{r Haversine Function}
Haversine <- function(lat1,lon1,lat2,lon2){
  # convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180
# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a))
# Earth's radius in kilometers
earth_radius <- 6378137
# Calculate the distance
distance_km <- (earth_radius * c)/1000
return(distance_km)
}
```

#5. Distance from New York to Auburn
```{r Auburn to NY}
AUNY <- Cities2 %>%
  dplyr::select(city1, city2, lat, long, lat2,lon2) %>%
  subset(city1 == c("New York"))
#subsetting so only the distance from new york to auburn is calculated
  
Haversine(AUNY$lat, AUNY$long, AUNY$lat2, AUNY$lon2)
#the haversine function above, using the subsetted data to calculate the distance

```

#6 and Bonus. Distances from Auburn to cities
```{r Auburn to Cities}

distances_km <- NULL #creates a null object so tidyverse will work
AUDistance <- Cities2 %>%
  group_by(city) %>% #grouping by city
  nest() %>% #nesting the data
  mutate(distances_km= map(data, ~Haversine(.$lat,.$long,.$lat2,.$lon2))) %>% #creates a new column after mapping the data and doing the haversine function
  unnest(c(data,distances_km)) %>% #unnests data and distances
  ungroup(city) %>% #ungroups city so we can use the select function
  dplyr::select(city1, city2, distances_km) #selects three columns we want in the resulting dataframe (removes excess garbage)
 AUDistance
```




